Optimal Fitting of Planar Curves to Prescribed Constraints

### 2013-08-29 idp/documentation/paper.tex
reorganized proposed solution section

			We think that in order to get the best usability, using a good specification language which does not rely on implementation details has to be the top priority. Since it is not immediately clear which specification language is the best, the first step is to move away from the traditional explicit curve construction, and instead use a process that enables the creation of curves based on some specification and some fairness measure.

@@

			Figure \ref{figure:specification-based_curves} shows such a process. Note that many specifics are still left undetermined in this approach since the previous usability analysis didn't give a complete answer as to what the optimal curve design tool looks like. It has yet to be determined what the specification language and fairness measure should be, whether to do segmentation (spline approach) or use a single parametric curve, or if the curve should be determined by numerical optimization or by explicit construction.

		\subsection{Parametric Plane Curves}
		\label{section:parametric_plane_curves}

@@

		\subsection{Continuity}
		\label{section:continuity}

@@

			After having established the general approach to curve design tools that the usability analysis seems to suggest in section \ref{section:specification-based_curves}, it's time to decide on the specifics in order to create an instance of this model.

			The goal of this project is to figure out which specification language and fairness measure yield the best usability in the resulting curve design tool. In order to achieve this goal, we decided to use nonlinear optimization on polynomial splines, that is, the curves are splines built from parametric polynomial curves whose parameters are determined using nonlinear optimization on some objective function determined by the specification and the fairness measure. This approach allows us to conveniently experiment with specification languages and fairness measures in order to assess their usability.

@@

				% TODO: add citations

@@

@@

				Having chosen a specification language and a fairness measure, it is time to state the optimization problem. Since polynomials of very high degree usually don't behave very well, an approach where multiple polynomial parametric curves are connected to a spline seems advisable. This way, the optimization domain is a vector consisting of all the coefficients of all the polynomials that make up the spline. In the following, we will denote the \(i\)th coefficient of the \(k\)th polynomial by \(c_{k,i}\), and the vector of all coefficients by \(c\).

				% TODO: figure out proper names

@@

				Since we chose to give the position of a specification item as a fraction of the curve's total arc length, we need the function mapping the covered arc length to the corresponding parameter values in order to state the error terms for the specification items. Unfortunately, this function isn't very well suited to be used as a term in a smooth opimization problem even for simple curves like polynomials. Also, since the function isn't fixed and varies with the coefficients, there is no way to tell on which polynomial segment each specification item will end up, something that is difficult to state in a smooth optimization problem. Another issue is the need to incorporate the specification of the total arc length of the curve into the optimization goals, which is also related to the covered arc length function. In order to address these problems, the following error term is used as part of the objective function:
				\begin{equation*}
					\apply{e_s}{c} = \int_{0}^{1}\left(\apply{s_c}{t} - l^*\right)^2\;\mathrm{d}t
				\end{equation*}
				For polynomial splines with sufficient degree and segment count, this error term will be brought close to zero, making the speed approximately constant with value \(l^*\). This, in turn, makes the covered arc length function approximately linear (\(\apply{l_c}{t} \approx l^*t\)) which causes the inverse of the covered arc length function \(l_c^{-1}\) to be approximately linear as well, facilitating the simple statement of the error terms for the specification items. It also causes the total arc length of the curve to be equal to \(l^*\).

				As mentioned in section \ref{section:specification_language}, we are not concerned with properties beyond the second derivative of the curve. Thus, we will only require \(\mathrm{G}^2\) continuity for the whole spline. While the polynomial curve segments are smooth (\(\mathrm{C}^\infty\)), the resulting spline in general is not. The segments have to be pieced together in such a way as to ensure a certain degree of smoothness. Thus, to ensure \(\mathrm{G}^2\) continuity for the whole curve, it is sufficient to ensure \(\mathrm{G}^2\) continuity at the segment connection points. Since the constant speed requirement introduced in the previous paragraph collapses \(\mathrm{C}^2\) continuity and \(\mathrm{G}^2\) continuity, ensuring \(\mathrm{C}^2\) continuity is equivalent, yet better suited for optimization. Thus, for each connection point (\(0 \leq k \leq n - 1\)), three error terms are added:
				\begin{equation*}
					\begin{aligned}
						\apply{e_a}{c} & = \apply{p_{c_k}}{1} - \apply{p_{c_{k+1}}}{0}\\
						\apply{e_b}{c} & = \apply{p'_{c_k}}{1} - \apply{p'_{c_{k+1}}}{0}\\
						\apply{e_c}{c} & = \apply{p''_{c_k}}{1} - \apply{p''_{c_{k+1}}}{0}
					\end{aligned}
				\end{equation*}

				% TODO: introduce constraints and objective model of ipopt

				% segmentation, separation of segmentation and specification, global positions on curve
				% give objective and constraints function
				% check code to see if anything was left out

>>>>

			We think that in order to get the best usability, using a good specification language which does not rely on implementation details has to be the top priority. Together with a good fairness measure, it will then lead to usable curve design tools. Since it is not immediately clear which specification language or which fairness measure is the best, the first step is to move away from the traditional explicit curve construction, and instead use a process that enables the creation of curves based on explicit specifications and a fairness measure.

@@

			Figure \ref{figure:specification-based_curves} shows such a process. Note that many specifics are still left undetermined in this approach since the previous usability analysis didn't give a complete answer as to what the optimal curve design tool looks like. It has yet to be determined what the specification language and fairness measure should be, whether to do segmentation (spline approach) or use a single curve, or if the curve should be determined by numerical optimization or by explicit construction.

		\subsection{Definitions}
		\label{section:definitions}

			TODO

			\subsubsection{Parametric Plane Curves}
			\label{section:parametric_plane_curves}

@@

			\subsubsection{Continuity}
			\label{section:continuity}

@@

		\subsection{Specification Language and Fairness Measure}
		\label{section:specification_language_fairness_measure}

			After using the results of the usability analysis to establish the general approach to curve design tools in section \ref{section:specification-based_curves}, it's time to decide on the specifics in order to create an instance of this model. The first steps are to choose both a specification language and a fairness measure.

@@

				% TODO: add citations
				% TODO: make clear that these terms are to be minimized

@@

			Having decided on a specification language and a fairness measure in section \ref{section:specification_language_fairness_measure}, the only thing missing from the process described in section \ref{section:specification-based_curves} is a way of turning specifications into actual curves. In order to achieve this goal, we decided to use nonlinear optimization on polynomial splines, that is, the curves are splines built from parametric polynomial curves whose parameters are determined using nonlinear optimization based on the specification and the fairness measure.

@@

				Nonlinear optimization problems are usually expressed using an objective function \(f : \mathbb{R}^n \rightarrow \mathbb{R}\), a constraint function \(g : \mathbb{R}^n \rightarrow \mathbb{R}^m\) and constraint bounds \(g_l, g_u \in \mathbb{R}^m\). The optimization problem is then given by the task of finding an \(x^* \in \mathbb{R}^n\) such that:
				\begin{equation*}
					\begin{aligned}
						x^* = \min_{x \in \mathbb{R}^n} \apply{f}{x}\\
						g_l \leq \apply{g}{x^*} \leq g_u
					\end{aligned}
				\end{equation*}
				This allows expression of a variety of optimization goals. The ones that are important for us are zero-error constraints (using the constraint function and bounds to specify that an error term is zero), upper bounds on errors (using the constraint function and bounds to specify an upper bound for an error term) and minimization goals (using the objective function to express that an error term should be minimized).

				In the following sections, we will introduce the optimization goals that enable us to use nonlinear optimization for building curves based on specifications using the specification language and fairness measure established in section \ref{section:specification_language_fairness_measure}. We will also state how they are expressed using constraint function and bounds as well as objective function.

			\subsubsection{Constant Speed}
			\label{section:constant_speed}

				Since we chose to give the position of a specification item as a fraction of the curve's total arc length, we need the function mapping the covered arc length to the corresponding parameter values in order to state the optimization goals for the specification items. Unfortunately, this function is not very well-suited to be used as a term in a smooth opimization problem even for simple curves like polynomials. Also, since the function isn't fixed and varies with the coefficients, there is no way to tell on which segment each specification item will end up, something that is difficult to state in a smooth optimization problem. Another issue is the need to incorporate the specification of the total arc length of the curve into the optimization goals, which is also related to the covered arc length function. In order to address these problems, the following error term is used:
				\begin{equation*}
					\apply{e_s}{c} = \max \left\{\left|\apply{s_c}{t} - l^*\right| \; \middle| \; t \in \left[0, 1\right]\right\}
				\end{equation*}
				For polynomial splines with sufficient degree and segment count, this term will be brought close to zero in the optimization, making the speed approximately constant with value \(l^*\). This, in turn, makes the covered arc length function approximately linear (\(\apply{l_c}{t} \approx l^*t\)) which causes the inverse of the covered arc length function \(l_c^{-1}\) to be approximately linear as well, facilitating the simple statement of the optimization goals for the specification items as well as fixing the segment for each specification item. It also causes the total arc length of the curve to be approximately equal to \(l^*\), as requested by the specification.

				Since this goal can only be fulfilled approximately by polynomial curves, it is impossible to state it as a zero-error constraint. It could however be stated as both an upper bound on errors and a minimization goal. Both approaches were implemented and tested, with the latter proving to yield a more stable optimization process.

				Since the error term given above is not well-suited for smooth optimization, the following approximation was finally used as part of the objective function:
				\begin{equation*}
					\apply{e_s}{c} = \int_{0}^{1}\left(\apply{s_c}{t} - l^*\right)^2\;\mathrm{d}t
				\end{equation*}

			\subsubsection{Continuity Connections}
			\label{section:continuity_connections}

				As mentioned in section \ref{section:specification_language}, we are not concerned with properties beyond the second derivative of the curve. Thus, we will only require \(\mathrm{G}^2\) continuity for the whole spline. While the polynomial curve segments are smooth (\(\mathrm{C}^\infty\)), the resulting spline in general is not. The segments have to be pieced together in such a way as to ensure a certain degree of smoothness. Thus, to ensure \(\mathrm{G}^2\) continuity for the whole curve, it is sufficient to ensure \(\mathrm{G}^2\) continuity at the segment connection points. Since the constant speed requirement introduced in the previous paragraph collapses \(\mathrm{C}^2\) continuity and \(\mathrm{G}^2\) continuity, ensuring \(\mathrm{C}^2\) continuity is equivalent, yet better suited for optimization. Thus, for each connection point (\(0 \leq k \leq n - 1\)), the following error terms are added:
				\begin{equation*}
					\begin{aligned}
						\apply{e_a}{c} & = \apply{p_{c_k}}{1} - \apply{p_{c_{k+1}}}{0}\\
						\apply{e_b}{c} & = \apply{p'_{c_k}}{1} - \apply{p'_{c_{k+1}}}{0}\\
						\apply{e_c}{c} & = \apply{p''_{c_k}}{1} - \apply{p''_{c_{k+1}}}{0}
					\end{aligned}
				\end{equation*}

			\subsubsection{Specification Items}
			\label{section:specification_items}

			\subsubsection{Fairness Error}
			\label{section:fairness_error}

			\subsubsection{Complete Formalization}
			\label{section:complete_formalization}

				Since polynomials of very high degree usually don't behave very well, an approach where multiple polynomial parametric curves are connected to a spline seems advisable. This way, the optimization domain is a vector consisting of all the coefficients of all the polynomials that make up the spline. In the following, we will denote the \(i\)th coefficient of the \(k\)th polynomial by \(c_{k,i}\), and the vector of all coefficients by \(c\).

				% TODO: proper explanation of how segments are put together to a spline (wrt. parameter t)
				% TODO: figure out proper names, use asterisks consistently
				% TODO: where to use squares in errors terms, absolute values, etc.

				The continuity connections need exhibit zero error, since it is not acceptable if the curve segments don't touch, have sharp corners or curvature discontinuities. The specification item error terms should also have zero error, since the user usually wants to specify exact curve behavior rather than indirectly influence its properties. It's clear that in many cases, the fairness error term will not be zero, and in fact, may be arbitrarily large, so the fairness error term is best stated as an optimization goal rather than a hard bound on the error.

@@

				% give objective and constraints function
				% check code to see if anything was left out
